Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys

### 2019-01-25 doc/article.tex
forging on

\begin{abstract}
Cadence wide-field imaging surveys 
in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
their absolute magnitudes, now make them competitive probes of the growth of structure.  The peculiar velocity power spectrum
is sensitive to $f\sigma_8$,
the product of the linear growth and amplitude of density perturbations.  Cross-correlation with synergistic galaxy surveys further constrains $f \sigma_8$ and the galaxy bias. 
Thus, in the next decade the peculiar velocities of
SNe~Ia will provide the growth of structure  in the local $z<0.3$ Universe as a powerful test of General Relativity and other models of gravity.
\end{abstract}

@@

From the above calculations, we find empirically that shot noise is important since higher number densities and lower intrinsic magnitude dispersion
significantly reduce $\gamma$ uncertainty, but that sample variance is also relevant  by the end of a ten-year survey.
Survey design is aided by understanding the circumstances in which  the surveys become sample variance limited.
First note that the  rest-frame volumetric SN~Ia rate (full selection efficiency) is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$;
after a few years the RSD measurement is sample variance limited.  The  contributions of $P_{vv}$ and
$\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
sare shown in  Figure~\ref{power:fig}.  For all of these survey scenarios, the transition between sample variance and shot-noise equality 
occurs at a $k$ squarely within the $k$-range that carries the most signal.

@@

>>>>

\begin{abstract}
Cadenced wide-field imaging surveys 
in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
their absolute magnitudes, now make them competitive probes of the growth of structure and hence of gravity.  The peculiar velocity power spectrum
is sensitive to $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
In the next decade the peculiar velocities of
SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ 
down to $0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
\end{abstract}

@@

From the above calculations, we find empirically that shot noise is important since higher number densities and lower intrinsic magnitude dispersion
significantly reduce $\gamma$ uncertainty, but that sample variance is also relevant  by the end of a ten-year survey.
Survey design is aided by understanding the circumstances in which  the surveys become sample variance limited.
The  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$;
after a few years the RSD measurement is sample variance limited.  The  contributions of $P_{vv}$ and
$\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
can be compared using  Figure~\ref{power:fig}.  For all of these survey scenarios, sample variance and shot-noise equality 
occurs at a $k$  within the range that carries the most signal.  However, the ten-year surveys do reach the sample variance
limit if   $k_{max} < 0.2$~$h$\,Mpc$^{-1}$.

@@

Modeling $k_{max}$ is a high as possible.
